Eye-tracking on the web: lessons learned from replicating 6 experiments

### 2021-06-30 manuscript.Rmd
minor

> here we will describe the analyses that correlate calibration quality with effect size at the individual (Josh H?)

@@

@@

@@

@@

The third study was a replication attempt of @Mann???.

@@

>>>>

> here we will describe the analyses that correlate calibration quality with effect size at the individual

@@

Participants for this study were recruiting using the website Prolific. Specifically, participants had to be older than 18 and fluent in English, but aside from that, there was no restriction on demographic; participation was also anonymous. The only technology-based restriction was that each participant had to have a working webcam; we were able to accommodate for different screen sizes during data analysis. We analyzed the data of 59 participants, a number that was limited by budget constraints, but still 2.5x larger than the original sample size of 24, as suggested by Simonsohn (2015). We ended up excluding the data of 1 participant whose eye tracking data seemed to be blank, rendering us unable to analyze it.

@@

After a participant began the study, they would encounter an alert that the study would use their webcam to track eye movements, then initiate Webgazer, a program that logs predicted eye positions. There were two possible conditions they could begin with—the free-viewing condition and the fixed-viewing condition—which they were randomly assigned to. For both conditions, participants began with the encoding phase before moving to the recall phase. After going through both conditions, participants were asked a few survey questions and the experiment ended.
For the encoding phase, participants were asked to remember the contents of the four quadrants of a grid, each with six distinct items themed around a certain category (we used humanoids, household objects, animals, and methods of transportation, as inspired by the example in Johansson & Johansson’s original experiment). To do so, each of the four quadrants was presented to the participant one at a time. First, a list of the items in the quadrant were shown, then the items in the actual quadrant were shown. For each item in a quadrant, an audio file would play, asking the participant to use their arrow keys to identify which direction each item was facing (every item was facing a distinct direction to allow for statements like “The chair is facing left” to be viable). After the participant identified the direction of each item, they would have an additional 30 seconds to study and remember the name and orientation of each item in the quadrant. Then, after repeating this for each quadrant, the participant was shown the full grid of 24 items (six per grid) and had sixty seconds to further encode the name and orientation of each item.
For the recall phase, participants had to respond to statements presented via audio files. Each statement was a true or false statement that fell into either an interobject or intraobject condition. Interobject statements were those that compared two different items in the grid (e.g. “The skeleton is to the left of the robot”), while intraobject statements were those that asked about the orientation of a single item (e.g. “The bus is facing right”). There were 48 total statements, with 24 interobject and 24 intraobject statements split evenly among the four quadrants. Participants were able to respond by pressing the ‘F’ key for false statements and ‘T’ for true ones.
The difference between the free-viewing and fixed-viewing conditions was in what a participant could see on screen during the recall phase. During fixed-viewing, participants could only see a small cross in the center of the screen, and were asked at the start of the recall phase to focus their vision on just the cross. During free-viewing, participants saw an empty screen and were allowed to look wherever they wanted, without any particular instruction telling them where to look. In both cases, the mouse was obscured from the screen, making it so that the only visual stimulus for either condition was the cross in the fixed-viewing one. After a participant finished the first condition, they then moved on to repeating the experiment under the second condition, this time with a new grid of images. After completing this second condition, the participant was finished, and was asked to answer a few survey questions (such as whether they wore glasses or encountered any distractions).
The methodology of this replication differed from Johansson & Johansson’s original study in two key ways. First, the original study included two more conditions that were omitted from this replication for efficiency concerns. Those two conditions involved prompting the participant to look to an area on screen that matched one of the original quadrants while they responded to each statement, with the key here being that sometimes the prompt was in the same quadrant as the information being recalled, while sometimes it wasn’t. We ended up not replicating this aspect of the study because it felt like too big of a task to program and playtest in the amount of time we had. The second major way this study differed from the original was that the original was conducted in- person and used an iView Red500 eye tracker, which was able to track eye movements incredibly closely. Due to obvious pandemic-related reasons, we conducted our replication online through Prolific, while using Webgazer, a software that tracks eye movements based on a participant’s webcam. There were a few other minor ways in which our study differed (we used different grids of objects and different instructions), but we tried to remain as faithful as possible in terms of variables like timing, the physical appearance of the grids, and the style of the images.

@@

> here we will describe the analyses that correlate calibration quality with effect size at the individual

@@

The third study was a replication attempt of @mannsVisualPairedcomparisonTask2000.

@@

> here we will describe the analyses that correlate calibration quality with effect size at the individual

### Effects of ROIs

> here we will describe how results change depending on the size of the ROIs (using the image vs the screen quadrant)

### 2021-06-30 manuscript.tex
minor

\begin{quote}
here we will describe the analyses that correlate calibration quality with effect size at the individual (Josh H?)
\end{quote}

@@

@@

@@

The third study was a replication attempt of (\textbf{Mann?})???.

@@

@@

>>>>

\begin{quote}
here we will describe the analyses that correlate calibration quality with effect size at the individual
\end{quote}

@@

After a participant began the study, they would encounter an alert that the study would use their webcam to track eye movements, then initiate Webgazer, a program that logs predicted eye positions. There were two possible conditions they could begin with---the free-viewing condition and the fixed-viewing condition---which they were randomly assigned to. For both conditions, participants began with the encoding phase before moving to the recall phase. After going through both conditions, participants were asked a few survey questions and the experiment ended.
For the encoding phase, participants were asked to remember the contents of the four quadrants of a grid, each with six distinct items themed around a certain category (we used humanoids, household objects, animals, and methods of transportation, as inspired by the example in Johansson \& Johansson's original experiment). To do so, each of the four quadrants was presented to the participant one at a time. First, a list of the items in the quadrant were shown, then the items in the actual quadrant were shown. For each item in a quadrant, an audio file would play, asking the participant to use their arrow keys to identify which direction each item was facing (every item was facing a distinct direction to allow for statements like ``The chair is facing left'' to be viable). After the participant identified the direction of each item, they would have an additional 30 seconds to study and remember the name and orientation of each item in the quadrant. Then, after repeating this for each quadrant, the participant was shown the full grid of 24 items (six per grid) and had sixty seconds to further encode the name and orientation of each item.
For the recall phase, participants had to respond to statements presented via audio files. Each statement was a true or false statement that fell into either an interobject or intraobject condition. Interobject statements were those that compared two different items in the grid (e.g.~``The skeleton is to the left of the robot''), while intraobject statements were those that asked about the orientation of a single item (e.g.~``The bus is facing right''). There were 48 total statements, with 24 interobject and 24 intraobject statements split evenly among the four quadrants. Participants were able to respond by pressing the `F' key for false statements and `T' for true ones.
The difference between the free-viewing and fixed-viewing conditions was in what a participant could see on screen during the recall phase. During fixed-viewing, participants could only see a small cross in the center of the screen, and were asked at the start of the recall phase to focus their vision on just the cross. During free-viewing, participants saw an empty screen and were allowed to look wherever they wanted, without any particular instruction telling them where to look. In both cases, the mouse was obscured from the screen, making it so that the only visual stimulus for either condition was the cross in the fixed-viewing one. After a participant finished the first condition, they then moved on to repeating the experiment under the second condition, this time with a new grid of images. After completing this second condition, the participant was finished, and was asked to answer a few survey questions (such as whether they wore glasses or encountered any distractions).
The methodology of this replication differed from Johansson \& Johansson's original study in two key ways. First, the original study included two more conditions that were omitted from this replication for efficiency concerns. Those two conditions involved prompting the participant to look to an area on screen that matched one of the original quadrants while they responded to each statement, with the key here being that sometimes the prompt was in the same quadrant as the information being recalled, while sometimes it wasn't. We ended up not replicating this aspect of the study because it felt like too big of a task to program and playtest in the amount of time we had. The second major way this study differed from the original was that the original was conducted in- person and used an iView Red500 eye tracker, which was able to track eye movements incredibly closely. Due to obvious pandemic-related reasons, we conducted our replication online through Prolific, while using Webgazer, a software that tracks eye movements based on a participant's webcam. There were a few other minor ways in which our study differed (we used different grids of objects and different instructions), but we tried to remain as faithful as possible in terms of variables like timing, the physical appearance of the grids, and the style of the images.

@@

\hypertarget{replication-1}{%
\subsubsection{Replication}\label{replication-1}}

\hypertarget{calibration-1}{%
\subsubsection{Calibration}\label{calibration-1}}

\begin{quote}
here we will describe the analyses that correlate calibration quality with effect size at the individual
\end{quote}

@@

The third study was a replication attempt of Manns, Stark, and Squire (2000).

@@

\hypertarget{replication-2}{%
\subsubsection{Replication}\label{replication-2}}

\hypertarget{calibration-2}{%
\subsubsection{Calibration}\label{calibration-2}}

\begin{quote}
here we will describe the analyses that correlate calibration quality with effect size at the individual
\end{quote}

\hypertarget{effects-of-rois}{%
\subsubsection{Effects of ROIs}\label{effects-of-rois}}

\begin{quote}
here we will describe how results change depending on the size of the ROIs (using the image vs the screen quadrant)
\end{quote}

@@

\leavevmode\hypertarget{ref-mannsVisualPairedcomparisonTask2000}{}%
Manns, J. R., Stark, C. E. L., \& Squire, L. R. (2000). The visual paired-comparison task as a measure of declarative memory. \emph{Proceedings of the National Academy of Sciences}, \emph{97}(22), 12375--12379. \url{https://doi.org/10.1073/pnas.220398097}
